---
title: "Data wragling - 2202"
author: "Yogesh"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data-wrangling exercises

```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(tibble)
library(fs)
```

```{r}
# Here goes the code... Note how this part does not get executed
```

```{r}
tibble(x = c(4, 3, 5, 1, 2))
tibble(x = c(4, 3, 5, 1, 2)) %>% filter(x > 2)
tibble(x = c(4, 3, 5, 1, 2)) %>% arrange(x)
tibble(x = c(4, 3, 5, 1, 2)) %>% arrange(desc(x))
tibble(x = c(4, 3, 5, 1, 2)) %>% arrange(desc(desc(x)))
tibble(x = c(4, 3, 5, 1, 2), y = c(2, 4, 3, 5, 1)) %>% select(x)
tibble(x = c(4, 3, 5, 1, 2), y = c(2, 4, 3, 5, 1)) %>% select(y)
tibble(x = c(4, 3, 5, 1, 2), y = c(2, 4, 3, 5, 1)) %>% select(-x)
tibble(x = c(4, 3, 5, 1, 2), y = c(2, 4, 3, 5, 1)) %>% select(-x, -y)
tibble(x = c(4, 3, 5, 1, 2)) %>% mutate(x_dbl = 2*x)
tibble(x = c(4, 3, 5, 1, 2)) %>% mutate(x_dbl = 2 * x, x_qdr = 2*x_dbl)
tibble(x = c(4, 3, 5, 1, 2)) %>% summarise(x_mu = mean(x))
tibble(x = c(4, 3, 5, 1, 2)) %>% summarise(x_max = max(x))
tibble(lbl = c("A", "A", "B", "B", "C"), x = c(4, NA, 5, 1, 2)) %>% group_by(lbl) %>% summarise(x_mu = mean(x), x_max = max(x))
tibble(lbl = c("A", "A", "B", "B", "C"), x = c(4, 3, 5, 1, 2)) %>% group_by(lbl) %>% summarise(n = n())
tibble(lbl = c("A", "A", "B", "B", "C"), x = c(4, 3, 5, 1, 2)) %>% count(lbl)
tibble(lbl = c("A", "A", "B", "B", "C"), x = c(4, 3, 5, 1, 2)) %>% summarise(n = n())
```

#Q1: 4,3,5
#Q2: 1,2,3,4,5
#Q3: 5,4,3,2,1,
#Q4: 1,2,3,4,5 (desc(desc(x)))
#Q5: 4,3,5,1,2
#Q6: 2,4,3,5,1
#Q7: 2,4,3,5,1
#Q8:
#Q9: x_dbl = 8,6, 10,2,4
#Q10: x_dbl and x_qdr
#Q11: we get a new tibble of the mean(x) which is one variable
#Q12: 5
#Q13: Grouip by A B C, and gives the mean according to groups, and max(x)
#Q14: Grouip by A B C and n = 5



### Working with Diabetes Data set
```{r}
diabetes_data <- read_csv(file = "data/diabetes.csv")
diabetes_data
View(diabetes_data)
```

#Q1: Observations - 403, Variables - 19
#Q2: The data is tidy
#Q3: chr - Character or strings, dbl - doubles or real number, int - integers


```{r}
diabetes_data<- diabetes_data %>% mutate(height_cm = round(height*2.54, 1),
                         weight_kg = round(weight/2.205, 1),
                         waist_cm = round(waist*2.54, 1),
                         hip_cm = round(hip*2.54, 1))
```

#Q4: 349
```{r}
diabetes_data_metric %>% 
  count(weight_m < 100)
```

#Q5: 53
```{r}
diabetes_data_metric %>% 
  count(weight_m > 100)
```

#Q6:51
```{r}
diabetes_data_metric %>% 
  filter(weight_m > 100) %>% 
  count(height_m < 160)
```
```{r}
diabetes_data_metric %>% 
  filter(weight_m > 100, height_m < 160) %>% 
  count()
```


#Q7: 4
```{r}
diabetes_data_metric %>% 
  filter(gender == "female", height_m > 180) %>% 
  count ()
```

#Q8: 51
```{r}
diabetes_data_metric %>% 
  filter(gender == "male", height_m > 180) %>% 
  count ()
```

#Q9: 95
```{r}
diabetes_data_metric %>% 
  filter(location == "Louisa", 
         gender == "female",
         age > 30) %>% 
count()
```

#Q10: 1
```{r}
diabetes_data_metric %>% 
  filter(location == "Buckingham", 
         gender == "male",
         age < 30,
         height_m > 190) %>% 
count()
```

#Q11: T2: Make a scatter plot of weight versus height and colour by gender for inhabitants of Louisa above the age of 40
```{r}
diabetes_data %>% 
  filter(location == "Louisa",
         age > 40) %>% 
  ggplot(mapping = aes(x = weight_kg,
                     y = height_cm,
                     color = gender)) +
  geom_point()
```

#Q12: T3: Make a boxplot of height versus location stratified on gender for people above the age of 50
```{r}
diabetes_data %>% 
  filter(age > 50) %>% 
  ggplot(mapping = aes(x = location,
                     y = height_cm,
                     fill = gender))+
  geom_boxplot()
```
#Q11: How old is the youngest person? 19
```{r}
diabetes_data %>% 
  arrange(age)
```
#Q12: 92
#Q13 :Of all the 20-year olds, what is the height of the tallest? 182.9
```{r}
diabetes_data %>% 
  filter(age == 20) %>% 
  arrange(height_cm)
```
#Q14: Of all the 20-year olds, what is the height of the shortest? 147.3
#Q15: Q15: How many columns (variables) starts_with a “b”? 4
#Q16: How many columns (variables) contains the word “eight”? 

```{r}
diabetes_data %>% 
select(starts_with("b"))
```

```{r}
diabetes_data %>% 
  select(contains("eight"))
```
#T5: Create a BMI_class variable


```{r}
diabetes_data <- diabetes_data %>%
  mutate(bmi = weight_kg/(height_cm/100)^2)
```


```{r}
tibble(x = rnorm(100)) %>% 
  mutate(trichotomised = case_when(x < -1 ~ "Less than -1",
                                   -1 <= x & x < 1 ~ "larger than or equal to -1 and smaller than 1",
                                   1 <= x ~ "Larger than or equal to 1"))
```

#T5
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(BMI_class = case_when(bmi < 18.5 ~ "underweight",
                               bmi >= 18.5 & bmi < 25 ~ "normal weight",
                               bmi >= 25 & bmi < 30 ~ "overweight",
                               bmi >= 30 & bmi < 35 ~ "obese",
                               bmi >= 35 & bmi < 40 ~ "severe obesity",
                               bmi >= 40 & bmi < 50 ~ "morbid obesity",
                               bmi >= 50 ~"super obese"
                               ))
```


#T6: Create a boxplot of hdl versus BMI_class

```{r}
ggplot(data = diabetes_data)+
  geom_boxplot(mapping = aes(x = reorder(BMI_class, hdl, FUN = median),
                                           y = hdl))
```
#Q17: People with diabetes who have highest hld are underweight.

#T7-T8
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(gender_num = case_when(gender == "male" ~ 1,
                            gender == "female" ~ 0))
```

```{r}
diabetes_data <- diabetes_data %>% 
  mutate(BFP = 1.39*bmi + 0.16*age - 10.34*gender_num -9)
```

```{r}
diabetes_data <- diabetes_data %>% 
  mutate(WHR = waist_cm/hip_cm)
```



#Q18:Which correlate better with BMI, WHR or BFP? (Hint: Is there a certain plot-type, which can visualise if two variables are correlated? Can you perhaps use an R-function to compute the “correlation coefficient”?) 

```{r}
library(patchwork)

p1 <- diabetes_data %>% 
  ggplot(mapping = aes(x = bmi,
                           y = WHR))+
  geom_jitter()+
  geom_smooth(method = lm)

p2 <- diabetes_data %>% 
  ggplot(mapping = aes(x = bmi,
                           y = BFP))+
  geom_jitter()+
  geom_smooth(method = lm)

```

```{r}
p1 + p2
```

#Q19:How many women and men are there in the data set?

```{r}
diabetes_data %>% 
  group_by(gender) %>% 
  summarise(n = n())
```
#Q20: How many women and men are there from Buckingham and Louise respectively in the data set?

```{r}
diabetes_data %>% 
  group_by(location) %>% 
  summarise(n = n())
```
#Q21: How many are in each of the BMI_class groups?
```{r}
diabetes_data %>% 
  group_by(BMI_class) %>% 
  summarize(n = n())
```
#Q22: Given the code below, explain the difference between A and B? Geom_bar does not need a y axis as it displays the number of observations of x on the y axis. Whereas, for geom_col the y axis needs to be specified
```{r}
# A
p3 <- diabetes_data %>%
  ggplot(aes(x = BMI_class)) +
  geom_bar()

# B
p4 <- diabetes_data %>%
  count(BMI_class) %>%
  ggplot(aes(x = BMI_class, y = n)) +
  geom_col()
```
```{r}
p3+p4
```
#Q23: What was the average age of the women living in Buckingham in the study?
```{r}
diabetes_data %>%
filter(gender == "female",
location == "Buckingham") %>%
summarise(mean_female_buckingham = mean(age))
```

#T9: T9: For each BMI_class group, calculate the average weight and associated standard deviation
```{r}
diabetes_data %>% 
  group_by(BMI_class) %>% 
  summarise(mean = mean(weight_kg), SD = sd(weight_kg, na.rm=FALSE))
```

#Q18:Which correlate better with BMI, WHR or BFP? (Hint: Is there a certain plot-type, which can visualise if two variables are correlated? Can you perhaps use an R-function to compute the “correlation coefficient”?) 

```{r}
diabetes_data %>%
select(bmi, BFP, WHR) %>%
drop_na() %>%
cor() %>%
round(4)
```


```{r}
ggscatter(data = diabetes_data, x = "bmi", y = "WHR",
                   add = "reg.line", conf.int = TRUE,
                   cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "red", fill = "lightgray"),
                   xlab = "Body mass index", ylab = "Waist-hip ratio")
```

```{r}
ggscatter(data = diabetes_data, x = "bmi", y = "BFP",
                   add = "reg.line", conf.int = TRUE,
                   cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "red", fill = "lightgray"),
                   xlab = "Body mass index", ylab = "Body fat percentage")
```


#Group assignment - another method to do it

```{r}
library(patchwork)
library(ggpubr)

bmi_whr <- ggscatter(data = diabetes_data,
x = "bmi",
y = "WHR",
add = "reg.line",
add.params = list(color = "red", fill = "lightgray"),
conf.int = TRUE)+
stat_cor(method = "pearson")

bmi_bfp <- ggscatter(data = diabetes_data,
x = "bmi",
y = "BFP",
add = "reg.line",
add.params = list(color = "red", fill = "lightgray"),
conf.int = TRUE)+
stat_cor(method = "pearson")

bmi_bfp+bmi_whr
```


```{r}
library(patchwork)
library(ggpubr)

bmi_whr1 <- ggscatter(data = diabetes_data, x = "bmi", y = "WHR",
                   add = "reg.line", conf.int = TRUE,
                   cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "red", fill = "lightgray"),
                   xlab = "Body mass index", ylab = "Waist-hip ratio")

bmi_bfp1 <- ggscatter(data = diabetes_data, x = "bmi", y = "BFP",
                   add = "reg.line", conf.int = TRUE,
                   cor.coef = TRUE, cor.method = "pearson",
          add.params = list(color = "red", fill = "lightgray"),
                   xlab = "Body mass index", ylab = "Body fat percentage")

bmi_bfp1+bmi_whr1
```
```{r}
sum(diabetes_data[["WHR"]])
```
```{r}
diabetes_data %>%
  summarise(my_mean = mean("BFP"))
```
```{r}
diabetes_data$bmi %>% 
  mean()
```
```{r}
diabetes_data <- diabetes_data %>% mutate_all(~replace(., is.na(.), 0))
View(diabetes_data)
```

```{r}
diabetes_data %>% 
  drop_na(BFP) %>% 
  summarise(mean(BFP))
```
```{r}
diabetes_data <- mutate(diabetes_data, BFP_ratio = BFP - mean(BFP))
```

```{r}
diabetes_data %>% 
  count(BMI_class) %>% 
  drop_na(BMI_class) %>% 
  mutate(pct = n/sum(n) * 100) %>% 
  ggplot(aes(x = BMI_class, y = pct))+
  geom_col()
```
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(BMI_class = factor(BMI_class,
                            levels = c("underweight", "normal weight", )
                            ))
```

























